---
title: "classwork"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(primer.data)
library(tidyverse) 
library(rstanarm)
library(skimr)
```


```{r}
glimpse(governors)
view(governors)
```

  #  What is the influence of age and gender on the expected longevity of political candidates?
  
  Wisdom:
  
  - Based in US -> governors, but we want to understand all types of political candidates
  - years of election from 1945-2011, if we are making predictions about the future and this data set is limited to nearly 10 years ago
  - people are living longer
  - not as many data points for women (increase of women's participation in recent decades)
  - includes election age, death age, and lived after
  
  population- question doesn't specify, but we want the population to be American governors
  validity
  representativeness- past governors (about the rows)
  
  justice- predictive or causal model
  - there's no causation without manipulation -> predictive model, we can't manipulate or randomly assign age or gender
  
  $lived\_after_{i} = \beta_0 + \beta_1 age + \beta_2 gender +\epsilon_{i}$
  
  