---
title: "classwork"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(primer.data)
library(tidyverse) 
library(rstanarm)
library(skimr)
```


```{r}
glimpse(governors)
view(governors)
```

  #  What is the influence of age and gender on the expected longevity of political candidates?
  
  Wisdom:
  
  - Based in US -> governors, but we want to understand all types of political candidates
  - years of election from 1945-2011, if we are making predictions about the future and this data set is limited to nearly 10 years ago
  - people are living longer
  - not as many data points for women (increase of women's participation in recent decades)
  - includes election age, death age, and lived after
  
  population- question doesn't specify, but we want the population to be American governors
  validity
  representativeness- past governors (about the rows)
  
  justice- predictive or causal model
  - there's no causation without manipulation -> predictive model, we can't manipulate or randomly assign age or gender
  
  $lived\_after_{i} = \beta_0 + \beta_1 election\_age_{i} + \beta_2 sex_{i} +\epsilon_{i}$
  
  we have to choose a specific sex for the equation
  $lived\_after_{i} = \beta_0 + \beta_2 male_{i} + \beta_1 election\_age_{i}  +\epsilon_{i}$ 
  
  **data generating mechanism- assumptions that we are making

```{r fitting the model}

fit_1 <-        stan_glm(formula = lived_after ~ sex + election_age,
                     data = governors,
                     refresh =0,
                     seed = 12)

print(fit_1)
```
   intercept = unknown female base age 60-72
   sexmale = 6.1 years higher 
  election_age = when comparing two candidates who differ by one year in election age then the older one will live .8 years less. slope = -.8
  
  
  
```{r fit interactive model}
fit_2 <-        stan_glm(formula = lived_after ~ sex + election_age + sex*election_age,
                     data = governors,
                     refresh = 0,
                     seed = 12)
```
$lived\_after_{i} = \beta_0 + \beta_1 election\_age_{i} + \beta_2 male_{i} +  \beta_3 male_{i}*election\_age_{i} + \epsilon_{i}$ 

```{r fit interactive model}
print(fit_2)
```
  
  intercept: lived after
  sexMale: 
  election_age: -0.5 to .7 - includes 0, so probably not helpful 
  the difference in longevity between two female candidates is .1 years
  sexMale:election_age - 
  
  easier for female candidates